
---
title: "Comparison of methods within 'easySdcTable'"
date: "`r Sys.Date()`"
author: "Øyvind Langsrud"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison of methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



## Introduction

This rmarkdown document tests the function `ProtectTable` in package `easySdcTable`. Thereby, some of the functionality in `sdcTable` is tested. 

Computing time as `milliseconds` are reported.

In practice the is done by redefining the function `ProtectTable` to be a functions that tests the functionality.

The total number of suppressed cells (`n_suppressed`) and the number of primary suppressed cells (`n_primary`) are counted.

A method based on the functionality in package `RegSDC` is used to reveal suppressed cells, if possible.  
The total number of revealed suppressed cells (`n_revealed`) and the number of revealed primary suppressed (`n_revealed_prim`) cells are counted.

In addition, if some zeroes are suppressed when zeros are not to be protected, this is also reported (`n_zeros_suppressed`).

All code to generate output is included in the document.  

Method `"Gauss"` is default in `ProtectTable` and is implemented independently of `sdcTable`.
Method `"GAUSS"` is the `sdcTable` implementation introduced in version 0.32.3.



```{r comment=NA, tidy = TRUE, results = "hide", message = FALSE}
library(easySdcTable)
library(Matrix)
library(RegSDC)
z3 <- EasyData("z3") 
```



```{r comment=NA}
ProtectTable <- function(..., maxN = 3, protectZeros = TRUE) {
  capture.output({a <- try(PTxyz(..., maxN = maxN, protectZeros = protectZeros, systemTime = TRUE))})
  if (class(a)[1] == "try-error") {
    return(invisible(as.character(a)))
  }
  k <- cbind(a$z, t(a$x) %*% cbind(a$y, 
       RegSDC::SuppressDec(a$x, a$z, a$y, rmse = pi/3, nRep = 3)))[which(is.na(a$z)), ]
  k <- as.matrix(k)
  k[, 3:5] <- RoundWhole(k[, 3:5])
  ind_revealed <- which(apply(k[, -(1:2)], 1, max) - apply(k[, -(1:2)], 1, min) == 0)
  revealed <- k[ind_revealed, , drop = FALSE]
  out <- c(milliseconds = 1000 * as.numeric(a$system_time[1]), 
    n_suppressed = sum(is.na(a$z)), 
    n_primary = sum(k[ ,2] <= maxN & k[ ,2] != protectZeros*(maxN+1)),
    n_revealed = length(ind_revealed), 
    n_revealed_prim = sum(revealed[ ,2] <= maxN & revealed[ ,2] != protectZeros*(maxN+1)))
  if (!protectZeros & any(k[ ,2] == 0))
    out["n_zeros_suppressed"] = sum(k[ ,2] == 0)
  return(out)
}
```


```{r comment=NA, tidy = TRUE}
for(i in 1:3) cat(capture.output(sessionInfo())[i],"\n")

unlist(sessionInfo()$otherPkgs$easySdcTable[c("Package", "Version", "Date")])

unlist(sessionInfo()$loadedOnly$sdcTable[c("Package", "Version", "Date")])

```


# Single table, 1200 cells

## Single table

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "Gauss", singletonMethod = "none")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "GAUSS", singletonMethod = "none")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "SIMPLEHEURISTIC")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "SIMPLEHEURISTIC_OLD")
```

## Single table, with singleton method

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "Gauss")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method = "GAUSS")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC", detectSingletons = TRUE)
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC_OLD", detectSingletons = TRUE)
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC", threshold = 3)
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC_OLD", threshold = 3)
```


## Single table, zeros protected

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "Gauss", singletonMethod = "none")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "GAUSS", singletonMethod = "none")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "SIMPLEHEURISTIC")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "SIMPLEHEURISTIC_OLD")
```

## Single table, zeros protected, with singleton method

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "Gauss")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method = "GAUSS")
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method ="SIMPLEHEURISTIC", threshold = 1)
ProtectTable(z3, c(1:2, 4:6), 7, protectZeros = TRUE, method ="SIMPLEHEURISTIC_OLD", threshold = 1)
```

# Linked tables, 1350 cells

## Linked tables

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "Gauss", singletonMethod = "none")
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "GAUSS", singletonMethod = "none")
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "SIMPLEHEURISTIC")
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "SIMPLEHEURISTIC_OLD")
```

## Linked tables, with singleton method

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "Gauss")
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method = "GAUSS")
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC", detectSingletons = TRUE)
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC_OLD", detectSingletons = TRUE)
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC",  threshold = 3)
ProtectTable(z3, 1:6, 7, protectZeros = FALSE, method ="SIMPLEHEURISTIC_OLD",  threshold = 3)
```


## Linked tables, zeros protected

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "Gauss", singletonMethod = "none")
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "GAUSS", singletonMethod = "none")
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "SIMPLEHEURISTIC")
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "SIMPLEHEURISTIC_OLD")
```

## Linked tables, zeros protected, with singleton method

```{r comment=NA, tidy = TRUE, message = FALSE}
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "Gauss")
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method = "GAUSS")
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method ="SIMPLEHEURISTIC", threshold = 1)
ProtectTable(z3, 1:6, 7, protectZeros = TRUE, method ="SIMPLEHEURISTIC_OLD", threshold = 1)
```



